---
title: "wjg397_exercise1"
author: "Walter Gray"
date: "2/7/2021"
output: html_document
---
## Data visualization: gas prices

### A) Box Plot  
```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
load(file = "Walter_Gray_Ex_1_Global_Env.RData")
ggplot(data=GasPrices) + 
  geom_boxplot(aes(x=Competitors, y=Price))
```  

Claim: Gas stations charge more if they lack direct competition in sight  
Conclusion: Although the price floor is roughly the same, gas stations without competitors in sight appear to charge substantially higher prices. This indicates the theory may hold water.

### B) Scatter Plot  
```{r, echo = FALSE}
ggplot(data = GasPrices, mapping = aes(x = Price, y = Income)) + 
  geom_point()
```  

Claim: The richer the area, the higher the gas price  
Conclusion: Prices do appear to trend upwards as income increases


### C) Bar Plot  
```{r, echo=FALSE}
ggplot(data = GasPrices) +
  geom_col (mapping = aes(x = Brand, y = Price)) 
```

Claim: Shell charges more than other brands  
Conclusion: It appears that Shell charges more than the other major brands listed in the dataset, but not more than brands that fall into the "other" category (which might be independent or smaller brands.)

### D) Faceted Histogram
```{r, echo=FALSE}
ggplot(data=GasPrices) + 
  geom_histogram(binwidth = .1, aes(x=Price), color = 'white') +
  facet_wrap(~ Stoplight)
```

Claim: Gas stations at stoplights charge more  
Conclusion: It does appear that gas stations at stoplights charge more (and there are more of them overall)

### E) Dealer's Choice: Boxplot
```{r, echo=FALSE}
ggplot(data=GasPrices) + 
  geom_boxplot(aes(x=Highway, y=Price))
```

Claim:Gas stations with direct highway access charge more  
Conclusion: Although the sample size is lower for gas stations by the highway, they do charge more.

## 2) Data visualization: a bike share network

### Plot A: Line Graph

```{r, echo=FALSE}
ggplot(data = bikeshare, 
       mapping = aes(x = hr, y = total)) +
  geom_line()
```

This plot shows total daily ridership (y axis, 'total') against the hour of the day (x axis, 'hr'). It's evident that ridership is low from the hours of midnight through 6AM and then spikes with morning communtes.There's a brief lull after the morning commute, followed by relatively steady afternoon ridership. Rides rise again and peak during evening rides (potentially return commutes or post-work leisure) before tapering off again as the night progresses.The lines connecting the measurement spikes imply that the most granular unit of time measurement here is an hour.  
One takeaway from this plot is that ridership corresponds to morning and early evening activity (presumably workers and students commuting) but a sustained mid-day ridership may also indicate tourists or other reasons for riding midday that are not well understood. 

### Plot B: Faceted Line Graph

```{r, echo=FALSE}
ggplot(data = bikeshare, 
       mapping = aes(x = hr, y = total)) +
  geom_line() +
  facet_wrap(~ workingday)
```

Like the previous plot, this shows total daily ridership (y axis, 'total') against the hour of the day (x axis, 'hr') with the additional insight of looking at non-workday ridership (presumably both weekends and holidays, noted on the lefthand plot "0") and workdays (represented by the righthand plot "1").  
Workday ridership is a much smoother progression from morning till night, and ridership continues later into the night than during workdays. For workday ridership, there are spikes in the morning and early evening likely related to work commutes. 

### Plot C: Faceted bar plot

```{r, echo=FALSE}
morning_bikeshare <- bikeshare %>%
  filter(hr==8)

ggplot(data = morning_bikeshare) +
  geom_col (mapping = aes(x = weathersit, y = total)) +
  facet_wrap(~ workingday)
```

These two bar charts show weekend and holiday ridership (The barchart '0' on the lefthand side) and workday ridership (the barchart '1' on the righthand side). Total rides are shown on the Y Axis and the severity of weather event is shown on the X axis, with 1 corresponding to 'Clear, Few clouds, Partly cloudy, Partly cloudy;' 2 corresponding to 'Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist;' 3 corresponding to 'Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds;' and 4 to 'Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog.'  
There are a number of insights that can be inferred from these charts. Workday ridership is significantly higher in general, and is most popular when weather conditions are clear. Some riders are willing to ride on workdays when it's overcast, but far fewer are likely to ride when it's raining or snowing. Amongst non-workday riders, most went out during clear conditions; some went out during overcast conditions. Few of the non-workday riders ventured out during snow or rain. Across both measured groups, no one rode at all during the worst weather conditions (heavy rain, hail, etc.).

## 3) Data visualization: flights at ABIA

### Which airlines had the most delays in 2008?

```{r, echo=FALSE, message=FALSE}
ABIA$total_delay <- ABIA$ArrDelay + ABIA$DepDelay
ggplot(data = ABIA) +
  geom_col (mapping = aes(x = UniqueCarrier, y = total_delay)) 
```

This first bar chart looks at the total amount of time delayed. To better understand the overall time spent on delays on a given flight, a new variable total_delay was created from combining arrival delays and departure delays. Negative values in the chart indicate the overall amount of time _saved_ from flights arriving ahead of schedule.  
The worst offender is Southwest Airlines (Unique Code is WN) followed by American Airlines (AA), Continental (CO), Jetblue (B6) and Mesa Airlines (YV, now defunct). Let's dig in a bit more to Southwest specifically. One caveat: the model removed 1601 rows due to incomplete data, so some airlines may be slightly over or understated in their time delays. 

### Is there seasonal variation in Southwest delays? 
```{r, echo=FALSE}
sw_flights <- ABIA %>% 
  filter(UniqueCarrier == "WN")
ggplot(data = sw_flights, 
       mapping = aes(x = Month, y = total_delay)) +
  geom_line() +
 scale_x_continuous(breaks=seq(0,12,1))
```

This graph attempts to answer the question of delay seasonality. Months of the year are displayed on the x-axis (where January = 1) and total delay time is listed on the y-axis.  
Delays were fairly consistent throughout the year, although there is a significant uptick in delays in December. Presumably holiday travel across both Southwest and other airlines is responsible for these delays. The time of year with the least delays is September, possibly because it is after summer vacation and lacks many holidays. The time of year with the highest likelihood to arrive _ahead_ of schedule is July. 

## 4) K-nearest neighbors
```{r, echo=FALSE}
library(rsample)  
library(caret)
library(modelr)
library(parallel)
library(mosaic)
library(readxl)
library(tidyverse)
library(ggplot2)
#Filter by trim
classA <- sclass %>% 
  filter(trim == "350")
# Make a train-test split
classA_split =  initial_split(classA, prop=0.9)
classA_train = training(classA_split)
classA_test  = testing(classA_split)
# KNN with K = 2
knn2 = knnreg(mileage ~ price, data=classA_train, k=2)
#rmse(knn2, classA_test)
#rmse is 13895.05 at k=2

#plot(price ~ mileage, data = classA_train)

p_test = ggplot(data = classA_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.2)
p_test
p_test + geom_line(aes(x = mileage, y = price), color='red', size=1.5)
```
```{r,echo=FALSE}
library(rsample)  
library(caret)
library(modelr)
library(parallel)
library(mosaic)
library(readxl)
library(tidyverse)
library(ggplot2)
#Filter by trim
classA <- sclass %>% 
  filter(trim == "350")
# Make a train-test split
classA_split =  initial_split(classA, prop=0.9)
classA_train = training(classA_split)
classA_test  = testing(classA_split)
# KNN with K = 10
knn10 = knnreg(mileage ~ price, data=classA_train, k=10)
#rmse(knn10, classA_test)
#rmse is 12793.45 at k=10

#plot(price ~ mileage, data = classA_train)

p_test = ggplot(data = classA_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.2)
p_test
p_test + geom_line(aes(x = mileage, y = price), color='red', size=1.5)
```
```{r}

```

